---
title: "Single Cell RNA-Seq : 03-Initial Refinement"
author: "Alexander G Foote"
date: "`r Sys.Date()`"
output:
    html_document:
      keep_md: TRUE
      toc: TRUE
---

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, error=FALSE, message=FALSE)
```

## Load packages

```{r include=FALSE}
library(Seurat)
library(DoubletFinder)
library(kableExtra) # format tables
library(ggplot2)  
library(biomaRt)
library(patchwork)
library(BiocManager)
library(tidyr) # create graphics
library(tidyverse)
library(viridis)   # accessible color palettes
library(glmGamPoi)  
set.seed(12345)
```

```{r}
saline1 <- readRDS("/Users/alexanderfoote/dev/Projs/R/scRNA_analysis_7dpi_IAV/rds/saline1.std_pipeline_singlets_unrefined_mt15_IAVseq.rds")
saline2 <- readRDS("/Users/alexanderfoote/dev/Projs/R/scRNA_analysis_7dpi_IAV/rds/saline2.std_pipeline_singlets_unrefined_mt15_IAVseq.rds")
IAV <- readRDS("/Users/alexanderfoote/dev/Projs/R/scRNA_analysis_7dpi_IAV/rds/IAV.std_pipeline_singlets_unrefined_mt15_IAVseq.rds")
```

## Identify clusters for refinement

For our initial dataset refinement, it is essential to filter out thyroid epithelial cells from our single-cell data of the larynx/trachea, as thyroid tissue is anatomically adjacent but not relevant to our analysis. We’ll first use Epcam to broadly identify epithelial cell populations, then specifically remove thyroid cells expressing markers Tg, Tpo, Folr1, and Pax8. This preliminary filtering step ensures our dataset accurately represents only the intended laryngeal and tracheal cell populations before proceeding to integration and further downstream refinement.

```{r}
#all major compartments
FeaturePlot(saline1, features = c("Epcam","Col1a2","Ptprc","Cldn5"), order=TRUE,pt.size=0.5) 
FeaturePlot(saline2, features = c("Epcam","Col1a2","Ptprc","Cldn5"), order=TRUE,pt.size=0.5) 
FeaturePlot(IAV, features = c("Epcam","Col1a2","Ptprc","Cldn5"), order=TRUE,pt.size=0.5) 
#Thyroid gland cell types
FeaturePlot(saline1, features = c("Epcam","Tg","Tpo","Folr1","Pax8"), order=TRUE,pt.size=0.5) 
VlnPlot(saline1,features=c("Epcam","Tg","Tpo","Folr1","Pax8"),pt.size=0)
FeaturePlot(saline2, features = c("Epcam","Tg","Foxe1","Tpo","Folr1","Pax8"), order=TRUE,pt.size=0.5) 
VlnPlot(saline2,features=c("Epcam","Tg","Tpo","Folr1","Pax8"),pt.size=0)
FeaturePlot(IAV, features = c("Epcam","Tg","Tpo","Folr1","Pax8"), order=TRUE,pt.size=0.5) 
VlnPlot(IAV,features=c("Epcam","Tg","Tpo","Folr1","Pax8"),pt.size=0)
```

## FindAllMarkers to confirm gene expression

```{r}
#setwd('/Users/alexanderfoote/R_Projects/scRNA_analysis_7dpi_postIAV/csv')
#Finding DE features (CLUSTER BIOMARKERS)
name <- "saline1-singlet_unrefined_allcompartments_IAVseq"
scRNA_saline1.markers <- FindAllMarkers(saline1, assay = "RNA", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
x=scRNA_saline1.markers %>% group_by(cluster) %>% top_n(n = 100, wt = avg_log2FC)
write.table(x, file=paste(name,"_Clustermarker.csv",sep=""), append = FALSE, quote = FALSE) 
name <- "saline2-singlet_unrefined_allcompartments_IAVseq"
scRNA_saline2.markers <- FindAllMarkers(saline2, assay = "RNA", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
x=scRNA_saline2.markers %>% group_by(cluster) %>% top_n(n = 100, wt = avg_log2FC)
write.table(x, file=paste(name,"_Clustermarker.csv",sep=""), append = FALSE, quote = FALSE) 
name <- "IAV-singlet_unrefined_allcompartments_IAVseq"
scRNA_IAV.markers <- FindAllMarkers(IAV, assay = "RNA", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
x=scRNA_IAV.markers %>% group_by(cluster) %>% top_n(n = 100, wt = avg_log2FC)
write.table(x, file=paste(name,"_Clustermarker.csv",sep=""), append = FALSE, quote = FALSE)
```

## Initial Refinement

Considering gene markers and feature plot, REFINE clusters to remove thyroid gland epithelial clusters

```{r}
saline1 <- subset(saline1,idents=c(0,1), invert = TRUE)
saline1 <- RunPCA(saline1, npcs = 100)
saline1 <- FindNeighbors(saline1, reduction = "pca", dims = 1:50)
saline1 <- FindClusters(saline1, resolution = 0.5) 
DimPlot(saline1, reduction = "umap", label=TRUE, pt.size = 0.5)

saline2 <- subset(saline2,idents=c(11), invert = TRUE)
saline2 <- RunPCA(saline2, npcs = 100)
saline2 <- FindNeighbors(saline2, reduction = "pca", dims = 1:50)
saline2 <- FindClusters(saline2, resolution = 0.5) 
DimPlot(saline2, reduction = "umap", label=TRUE, pt.size = 0.5)

IAV <- subset(IAV,idents=c(8), invert = TRUE)
IAV <- RunPCA(IAV, npcs = 100)
IAV <- FindNeighbors(IAV, reduction = "pca", dims = 1:50)
IAV <- FindClusters(IAV, resolution = 0.5) 
DimPlot(IAV, reduction = "umap", label=TRUE, pt.size = 0.5)
```

## Re-check Compartments and Feature Counts

```{r}
#all major compartments
VlnPlot(saline1,features=c("Epcam","Ptprc","Col1a1","Cldn5"),pt.size=0.0)
VlnPlot(saline1,features=c("nFeature_RNA","nCount_RNA","percent.mt"),pt.size=0)
VlnPlot(saline2,features=c("Epcam","Ptprc","Col1a1","Cldn5"),pt.size=0)
VlnPlot(saline2,features=c("nFeature_RNA","nCount_RNA","percent.mt"),pt.size=0)
VlnPlot(IAV,features=c("Epcam","Ptprc","Col1a1","Cldn5"),pt.size=0)
VlnPlot(IAV,features=c("nFeature_RNA","nCount_RNA","percent.mt"),pt.size=0)

#Thyroid gland cell types
VlnPlot(saline1,features=c("Epcam","Tg","Tpo","Folr1","Pax8"),pt.size=0)
VlnPlot(saline2,features=c("Epcam","Tg","Tpo","Folr1","Pax8"),pt.size=0)
VlnPlot(IAV,features=c("Epcam","Tg","Tpo","Folr1","Pax8"),pt.size=0)
```

### Save object

```{r}
#setwd('/Users/alexanderfoote/R_Projects/scRNA_analysis_7dpi_postIAV/rds')
saveRDS(saline1, "saline1.std_pipeline_singlets_refined_mt15_IAVseq.rds")
saveRDS(saline2, "saline2.std_pipeline_singlets_refined_mt15_IAVseq.rds")
saveRDS(IAV, "IAV.std_pipeline_singlets_refined_mt15_IAVseq.rds")
```

### Session information

```{r}
sessionInfo()
```
